\documentclass[12in]{article}

\usepackage{xcolor}
\usepackage{natbib}
% \usepackage{apalike}
\usepackage[colorlinks=true]{hyperref}
\usepackage{tikz}
\usepackage{caption}
% \usepackage{subcaption}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{subfig}

\linenumbers
\setpagewiselinenumbers

\usetikzlibrary{positioning,arrows}
\usetikzlibrary{shapes.geometric}

\newcommand{\referencessection}{\section*{References}}
% \setlength{\parindent}{0pt}

\title{Intention to submit document for the
\href{https://www.ukri.org/opportunity/work-with-us-researchers-bbsrc-nsf-bio-lead-agency-2024/}{Work
with US researchers BBSRC-NSF/BIO lead agency 2024} funding
opportunity\\\vspace{.5in}Enabling Naturalistic, Long-Duration and Continual
Neuroscience Experimentation with Advanced Machine Learning}

\begin{document}

\tableofcontents

\pagebreak

\maketitle

\section{Summary}

{\color{red}
Word limit: 2
A summary is not required for this section, please write ‘N/A’ in the textbox.
Please still include a title for your project.
}

N/A

\pagebreak

\section{Core team}
\label{sec:coreTeam}

{\color{red}

List the key members of your team and assign them roles from the following:

\begin{itemize}
    \item project lead (PL)
    \item project co-lead (UK) (PcL)
    \item specialist
    \item professional enabling staff
    \item research and innovation associate
    \item technician
    \item researcher co-lead (RcL)
\end{itemize}

Only list one individual as project lead.

The core team section must only contain details of the UK applicants. The US
applicant information should be listed in the `US applicants' section.

Find out more about
\href{https://www.ukri.org/publications/roles-in-funding-applications/roles-in-funding-applications-eligibility-responsibilities-and-costings-guidance/}{UKRI’s
core team roles in funding applications}.

}

\input{coreTeam}

\section{Application questions}

\subsection{Research theme}

{\color{red}
Word limit: 5
Please state the research theme you are applying under. Choose one of the following research themes:

\begin{enumerate}
	\item biological informatics
	\item understanding host-microbe interactions
	\item synthetic cells and cellular systems
	\item synthetic microbial communities
\end{enumerate}
}

biological informatics

\subsection{Vision}

{\color{red}
Word limit: 500

What are you hoping to achieve with your proposed work?

What the assessors are looking for in your response

Your vision should clearly address:

\begin{itemize}

	\item one of the opportunity research themes (biological informatics,
understanding host-microbe interactions, synthetic cells and cellular systems
or synthetic microbial communities)

	\item the remit of the BBSRC and the NSF/BIO division associated with your
chosen research theme

\end{itemize}

References may be included within this section, but this will count towards your
word count.

Images are not required for this section.
}

\input{vision}

\subsection{Approach}

{\color{red}
Word limit: 500

How are you going to deliver your proposed work?

What the assessors are looking for in your response

Your approach should give an overview highlighting:

\begin{itemize}

	\item a clear description of the objectives and methodology for the
proposed work, including the contributions of the UK and US teams

	\item the potential outputs and outcomes of the proposed work

\end{itemize}

References may be included within this section, but this will count towards your
word count.

Images are not required for this section.
}

\input{approach}

\subsection{US applicants}

{\color{red}

Word limit: 200

Please provide the following details of the US applicants on this application:

\begin{enumerate}
	\item name
	\item institute
	\item job title
	\item role in project (for example, project lead or project co-lead)
	\item email address
\end{enumerate}

Please also indicate who the lead US applicant will be.

NSF will use this information to confirm applicant eligibility.

Please do not include details of US applicants in the ‘Core team’ section.

}

\subsection{Resources}

{\color{red}

Word limit: 200

Please provide the following:

\begin{itemize}

	\item overall estimates for costings and staffing full time equivalent (FTE) for both
the UK and US components

	\item clear separation of UK and US costings, in pounds sterling and US dollars
(USD) respectively

\end{itemize}

The overall budget should be below the maximum £2 million combined funder
contribution

If there is more than one UK or US team associated with the application, please
combine their estimates together.

A detailed calculation and breakdown of resources is not required at this stage,
nor is a justification of costs.

The following is an example of how this might look.

UK Resources:

Total cost estimate: £600,000

Research council contribution: £480,000

0.2 FTE time, 1.0 FTE PDRA, 0.5 FTE technician

US Resources:

Total cost estimate: \$300,000

1.0 FTE PDRA or 1.0 FTE doctoral researcher

Total funder contribution estimate:

£716,475 (£480,000 + £236,475 (\$300,000 at exchange rate 0.79))

}

\referencessection
\addcontentsline{toc}{section}{References}

\bibliographystyle{plain}
\bibliography{foraging,linearDynamicalSystems,stateSpaceModels,machineLearning,epilepsy,latentsVariablesModels,deepNeuralNets,monitoringBehavior,dataSharing,misc,naturalExperimentation,neuroEthology,bci,neuralDecoders,replay,optogenetics,tracking,benchmarks,physiology,aeon,others,spikeSorting}

\appendix

\section{More details about the vision}

\input{visionLong}

\end{document}
