
Few methods exist today for online spike sorting with high-channel-count
probes, as Neuropixels. However, online spike sorting is an active are of
research and new methods may appear soon. We will evaluate online spike sorting
methods on NaLoDuCo foraging recordings collected at the SWC and on NaLoDuCo
odour learning recordings collected at the AIND.

Two current methods to include in this comparison are
FAST and RT-Sort.
%
FAST~\citep{dhawaleEtAl17}, described in Section~\ref{sec:offlineSpikeSorting},
is a real-time spike sorting method, but its authors only used it offline. We
will attempt an efficient implemenation suitable for real-time spike sorting.
%
RT-Sort~\citep{vanDerMolenEtAl24} is a real-time spike detection and
sorting algorithm developed to enable low-latency processing of high-density
electrophysiological recordings, such as those obtained with Neuropixels
probes.

RT-Sort operates in two stages. First, it uses a lightweight
convolutional neural network (CNN) to continuously detect spikes in
streaming multichannel voltage recordings.
%
Second, RT-Sort performs spike sorting based on the spatiotemporal
propagation patterns of the detected spikes. Specifically, it analyzes the
timing of voltage extrema across electrodes to characterize the propagation of
the spike waveform through the neural tissue. These features are matched to an
evolving library of unit templates, enabling robust and online classification
of spikes to neurons. Templates are updated dynamically as new spikes are
detected, allowing the system to adapt to waveform drift and changing recording
conditions.
%
It also offers robustness to non-stationarities such as
electrode drift and waveform shape changes, which are common in long-duration
recordings.

The above methods have been tested on short duration experiments. We will
bechmark their performance in NaLoDuCo recordings and, if necessary, we will
adapt them to efficiently process non-stationarities in NaLoDuCo
experimentation.

\subsubsubsection{Outputs}

\begin{enumerate}

    \item Repository with adaptions of online spike sorting methods to process
    NaLoDuCo recordings and benchmarking of these adaptations.

\end{enumerate}

