
\begin{frame}
    \frametitle{Deliverables}

    \begin{enumerate}

        \item software, hardware and papers demonstrating applications of
            intelligent experimental control in neuroscience experimentation.

    \end{enumerate}

\end{frame}

\begin{frame}
    \frametitle{Previous work}

    \begin{itemize}

        \item Bonsai is an excellent software for close-loop
            neuroscience experimental control.

        \item funded by a BBSRC
            award\footnote{\url{https://gow.bbsrc.ukri.org/grants/AwardDetails.aspx?FundingReference=BB\%2FW019132\%2F1}}
            we are integrating machine learning methods into Bonsai.

    \end{itemize}

\end{frame}


\begin{frame}
    \frametitle{Future work}

    \begin{itemize}

        \item Bonsai has been used for close-loop control using direct
            observations. With the advanced machine learning methods that we have
            added to Bonsai, we can know infer subtle patterns in behavioural and
            neural recordings not visible to the naked eye. Now we will find
            applications where these patterns can be used for experimental control.
            Fortunately, we work at the SWC, where such applications abound.

        \item it would be very useful to perform control based on patterns inferred
            from spiking activity recorded by Neuropixels probes. To do so, we need
            online spike sorting methods for Neuropixels probes. These methods are
            underdeveloped and, if necessary, we will need to develop one.

    \end{itemize}

\end{frame}

