\subsection{Deliverables}

\begin{enumerate}

    \item hardware specifications for recordings of behaviour and neural
        activity used at the
        SWC/GCNU\footnote{\url{https://sainsburywellcomecentre.github.io/aeon\_docs/reference/hardware.html}}
        and at the AIND for head-fixed
        foraging\footnote{\url{https://www.allenneuraldynamics.org/platforms/behavior}}
        and freely moving odour exploration.

    \item software for managing long-duration recordings (e.g., data storage, data
        indexing).

    \item software for online/offline quality control.

    \item software for creating alerts.

    \item software for online (behavioral and neural) data visualisation.

    \item software for online (behavioral and neural) data analysis.

\end{enumerate}

\subsection{Previous work}

\begin{itemize}

    \item the SWC has performed foraging experiments

    \begin{itemize}

        \item lasting xx weeks and recording behaviour only
        \item lasting yy weeks and recording behaviour and electrophysiology
        \item data is stored in files and in a MySQL database

    \end{itemize}

    \item the AIND has performed foraging experiments in head-fixed mice (ask Josh, Saskia or David about the status of this project).

    \item the AIND is setting up the odour exploration experiments.

    \item  items 1--2: above have been completed for the SWC foraging experiments

    \item item 3: the SWC has developed several quality control for behavior.

    \item item 4: alerts have been developed for behaviour.

    \item  item 5: the SWC has developed some online behavioural data
        visualisation tools in Bonsai.

    \item  item 6: funded by BBSRC, we have integrated into Bonsai tools for online data
  analysis:

    \begin{itemize}

        \item estimate kinematics of mice

        \item estimate kinematic states of mice using Hidden Markov Models

        \item clusterless point-process decoder of mice position and replay from spikes

    \end{itemize}

    \item  disseminated documentation of hardware used at the AIND to perform
        VR and head-fixed foraging
        exeperiments\footnote{\url{https://www.allenneuraldynamics.org/platforms/behavior}}.

    \item  disseminated documentation on software used at the SWC to control
        NaLoDuCo foraging experiments (see
        \href{https://github.com/SainsburyWellcomeCentre/aeon_mecha.}{repo})

    \item  disseminated documentation on machine learning methods integrated
        into Bonsai for analyzing behavioral and neural tie series in real time
        (see \href{https://bonsai-rx.org/machinelearning/.}{repo})

\end{itemize}

\subsection{Future work}

\begin{itemize}

    \item item 3: at the SWC and at the AIND we have developed several tools for offline quality control. We next need to build online versions of them.

    \item item 4: develop more software for data visualisation.

    \item items 5 and 6: develop more software for online data analysis.

    \begin{itemize}

        \item online estimate of latent variables from Neuropixels recordings.

        \item online estimate of RL models.

    \end{itemize}

    \item help Dr.~Carl Schoonover (AIND) use hardware and software developed at the SWC/GCNU to create his olfaction NaLoDuCo experiments.

\end{itemize}
