
\begin{instruction}
Word limit: 1,650

Why are you the right individual or team to successfully deliver the proposed
work?

What the assessors are looking for in your response

Please ensure the current job titles of the core team members are included
here to ensure eligibility can be established for the core team roles assigned.
Find out more about
\href{https://www.ukri.org/publications/roles-in-funding-applications/roles-in-funding-applications-eligibility-responsibilities-and-costings-guidance/}{UKRI’s
core team roles in funding applications} and our
\href{https://www.ukri.org/councils/bbsrc/guidance-for-applicants/check-if-youre-eligible-for-funding/applicants-and-co-applicants/}{eligibility
guidance}.

Evidence of how you, and if relevant your team, have:

\begin{itemize}

    \item the relevant experience (appropriate to career stage) to deliver the proposed
work

    \item the right balance of skills and expertise to cover the proposed work

    \item the appropriate leadership and management skills to deliver the work and
your approach to develop others

    \item contributed to developing a positive research environment and wider
community

\end{itemize}

You may demonstrate elements of your responses in visual form if relevant.

Further details are provided in the Funding Service.

The word limit for this section is 1,650 words: 1,150 words to be used for R4RI
modules (including references) and, if necessary, a further 500 words for
Additions.

Use the Résumé for Research and Innovation (R4RI) format to showcase the range
of relevant skills you and, if relevant, your team (project and project
co-leads, researchers, technicians, specialists, partners and so on) have and
how this will help deliver the proposed work. You can include individuals’
specific achievements but only choose past contributions that best evidence
their ability to deliver this work.

Complete this section using the R4RI module headings listed. Use each
heading once and include a response for the whole team, see the UKRI
guidance on R4RI. You should consider how to balance your answer, and
emphasise where appropriate the key skills each team member brings:

\begin{itemize}

    \item contributions to the generation of new ideas, tools, methodologies, or
knowledge

    \item the development of others and maintenance of effective working relationships

    \item contributions to the wider research and innovation community

    \item contributions to broader research or innovation users and audiences and
towards wider societal benefit

\end{itemize}

Additions

Provide any further details relevant to your application. This section is
optional and can be up to 500 words. You should not use it to describe
additional skills, experiences, or outputs, but you can use it to describe any
factors that provide context for the rest of your R4RI (for example, details of
career breaks if you wish to disclose them).

Complete this as a narrative. Do not format it like a CV.

References may be included within this section.

The roles in funding applications policy has descriptions of the different
project roles.

\end{instruction}
