
\begin{frame}
    \frametitle{Deliverables}

    \begin{enumerate}

        \item web-based dashboard for \textbf{online} experiment monitoring.

        \item web-based dashboard for \textbf{online} data analysis and
        visualisation of its results.

        \item web-based dashboard for \textbf{offline} visualisation of
            NaLoDuCo behavioural and neural recordings on DANDI.

            These dashboards should allow users to efficiently \textbf{visually
            explore} very large behavioural, neural and data analysis datasets.
            For instance, users should be able to quickly visualise videos of
            mice when they were next to any location of the arena, at any time
            of the day, and in any behavioural state (as inferred by a Hidden
            Markov Model). These efficient data explorations will require
            sophisticated data indexing schemes.

        \item web-based dashboard for \textbf{offline} visualisation of data
            analysis results on DANDI.

    \end{enumerate}

\end{frame}

\begin{frame}
    \frametitle{Previous work}

    \begin{itemize}

        \item Neurosift
            (\href{https://github.com/flatironinstitute/neurosift}{repo},
            \href{https://joss.theoj.org/papers/10.21105/joss.06590}{paper}) allows to
        visualise shorter-duration behavioural and neural recordings in DANDI.

        \item Dendro (\href{https://github.com/magland/dendro}{repo}) allows to
        perform analysis on the cloud and visualise the results of such analysis

        \item the SWC has developed visualisations in Bonsai for experimental
        monitoring.

        \item offline and precomputed visualisations developed at the SWC, of some
        experimental variables (e.g., total distance that animals have moved the
        wheel, proportion of time animals spent on the nest, corridor or food
        patches).

        \item offline and precomputed visualisations developed at the AIND for
        shorter-duration experiments.

        \item offline and precomputed visualisations from IBL for short-duration
        experiments.

    \end{itemize}

\end{frame}


\begin{frame}
    \frametitle{Future work}

    \begin{itemize}

        \item create web-based dashboard to monitor NaLoDuCo experiments.

        \item create web-based dashboard for \textbf{online} data analysis and
        visualisation of its results.

        \item create web-based dashboard for \textbf{offline} visualisation of NaLoDuCo
        behavioural and neural recordings on DANDI.

        Neurosift has been designed to visualise relatively short duration
        datasets. We will extend it with data pyramids (e.g.,
        \url{https://github.com/carbonplan/ndpyramid}) to enable it to operate on
        long-duration recordings.

        \item web-based dashboard for \textbf{offline} visualisation of data analysis
        results on DANDI.

        Dendro has been used to analyse on the cloud relatively small datasets. We
        will extend Dendro to analyse very large datasets and visualise the results
        of these analysis.

    \end{itemize}

\end{frame}
