\documentclass{beamer}

\usetheme{Berkeley} % Sidebar theme
% \setbeamertemplate{headline}{} % Removes title from sidebar
% \setbeamertemplate{sidebar left}{\insertsectionnavigation{.5\textwidth}{\hskip-1em}{}} % Removes author info

\usepackage{verbatim}
\usepackage{xcolor}
% Define custom colors (optional)
\definecolor{myblue}{RGB}{0, 0, 200} % Dark blue
\definecolor{myred}{RGB}{200, 0, 0}  % Dark red

% Set link colors
\hypersetup{
    colorlinks=true,
%     linkcolor=myblue,    % Color for internal links (e.g., table of contents)
    urlcolor=myred,      % Color for URLs
    citecolor=green      % Color for citations (if using BibTeX)
}

%Information to be included in the title page:
\title{Objectives and Specific Aims}
\author[]{Joaqu\'{i}n Rapela and Maneesh Sahani}
\institute{Gatsby Computational Neuroscience Unit}
\date{\today}

\AtBeginSection[]
{
 \begin{frame}<beamer>
 \frametitle{Contents}
 \tableofcontents[currentsection]
 \end{frame}
}

\begin{document}

\frame{\titlepage}

\begin{comment}
\begin{frame}
    \frametitle{Contents}

    \tableofcontents

\end{frame}
\end{comment}


\section{Objectives}

\begin{frame}
    \frametitle{Objectives}

    \begin{enumerate}

        \item \textbf{Co-develop a new type of neuroscience experimentation}
            that enables animals to express ethologically relevant behaviours
            over extended periods in naturalistic environments, while capturing
            high-precision behavioural and neural activity measurements.


        \item \textbf{Create a global resource to support research groups worldwide
            in adopting this new experimental framework}.

        \item \textbf{Enable global access to the data generated by these
            experiments} by developing web-based platforms for data access,
            visualisation, and analysis, while also allowing users to conduct
            custom analysis.

    \end{enumerate}

\end{frame}

\section{Specific Aims}

\begin{frame}
    \frametitle{Specific Aims}

    \begin{enumerate}

        \item \textbf{Develop software for neural and behavioural time-series
            visual exploration}, operating on continuous and epoched data.

        \item \textbf{Design efficient statistical data analysis methods} to
            process long-duration and non-stationary time-series at scale.

        \item \textbf{Develop batch and online spike sorting methods} to
            process weeks- to month-long continual electrophysiology recordings
            and support real-time machine learning inferences.

        \item \textbf{Build a framework for real-time machine learning algorithms} to
            enable intelligent neural manipulations.

    \end{enumerate}

\end{frame}

\begin{frame}
    \frametitle{Efficient data statistical analysis methods: challenges}

    \small
    \begin{description}

        \item[long-duration] most conventional machine learning method
            implementations for neural data analysis will not scale to datasets
            of the size generated in NaLoDuCo experiments. For example, (with
            classical methods) it will not be possible to invert a data matrix
            from a month-long experiment to perform batch linear regression.

            Possible solutions:

            \begin{enumerate}

                \item \textbf{distributed computing} some algorithms, like
                    batch linear regression, can be efficiently distributed
                    across several computers and be used in very large
                    datasets.
                    %
                    However, not all algorithms parallelise well, due to
                    sequential dependencies or high-communication costs, and
                    can be distributed efficiently.


                \item \textbf{online methods} some batch methods, like batch
                    linear regression,  have equivalent online methods, like
                    online Bayesian linear regression, which can process
                    infinite data streams.

            \end{enumerate}

        \item[non-stationarity] alternatives

            \begin{enumerate}
                \item detect performance degradation $\rightarrow$ retrain

                \item use non-stationary methods
            \end{enumerate}

    \end{description}
    \normalsize

\end{frame}

\begin{frame}
    \frametitle{Advanced statistical data analysis methods}

    \small
For behavioural data, we will investigate methods to:

\begin{description}

    \item[track multiple body parts of animals] (e.g.,
        [\href{https://pubmed.ncbi.nlm.nih.gov/30127430/}{7}] and a
        switching-linear-dynamical method using RFIDs that we will develop),

    \item[infer kinematics of foraging mice] (e.g.,
        [\href{https://github.com/joacorapela/lds\_python}{8},\href{https://www.cambridge.org/core/books/fundamentals-of-object-tracking/A543B0EA12957B353BE4B5D0602EE945}{9}]),

    \item[segment behaviour into discrete states] (e.g.,
        [\href{https://pubmed.ncbi.nlm.nih.gov/26687221/}{10}]
        and a hierarchical HMM that we will develop),

    \item[infer rules that govern mice behaviour from behavioural
        observations only] (i.e., policy inference) (e.g.,
        [\href{https://arxiv.org/abs/2311.13870v2}{11}]).

\end{description}

For neural data, we will investigate methods to:

\begin{description}

    \item[estimate low-dimensional continual representations of neural
        activity]
        (i.e., latents inference) (e.g.,
        [\href{https://papers.nips.cc/paper_files/paper/2011/hash/7143d7fbadfa4693b9eec507d9d37443-Abstract.html}{12}]),

    \item[segment neural activity into discrete states] (e.g.,
        [\href{https://pubmed.ncbi.nlm.nih.gov/21299424/}{13}]),

    \item[decode environment variables from neural activity] (e.g.,
        [\href{https://pubmed.ncbi.nlm.nih.gov/25973549/}{14}]).

\end{description}
\end{frame}

\section{Use Case}

\begin{frame}
    \frametitle{Use Case}

    \only<1>{
        A user begins by \textbf{visualising continuous behavioural measurements},
        such as the kinematics (speed and acceleration) of a mouse during a
        months-long experiment measured by the IMU of an ONIX probe.

        Next, she \textbf{examines the results of a machine learning analysis},
        such as behavioural states inferred by a Switching Hidden Markov Model (SHMM)
        using kinematic data.

        The SHMM was initially trained on the first two hours of the experiment
        and was periodically retrained to \textbf{adapt to non-stationarities},
        such as sensor fluctuations, changes in motivation, fatigue, or learning.
    }

    \only<2>{
        She then \textbf{visualises epoched data}, such as SHMM states aligned to
        key events—e.g., the onset of a foraging bout in the richer patch at a
        specific time of day.

        Curious about the neural basis of these SHMM states, she checks the neural
        recordings but realises they have not yet been spike-sorted. She runs our
        \textbf{offline spike sorting method}, developed for very long continuous
        recordings, and performs quality control on its results.

        Based on quality metrics, she detects drift in the recorded signal,
        adjusts the drift correction parameters, and reruns the sorting algorithm.
    }

    \only<3>{
		After examining the sorted spikes from a large neural population,  she
        finds it challenging to interpret activity across so many neurons.  So she
        decides to summarise the population spiking activity by \textbf{estimating
        continuous latent variables}.

        She then returns to the behavioural visualisation software, integrates
        machine learning indices corresponding to these latent variables, and
        \textbf{visualises behavioural data aligned to the newly estimated latents}.
    }

    \only<4>{
        Through these explorations, she hypothesises that a peak in a neural latent
        variable from the prefrontal cortex signals the moment when mice decide
        to begin a foraging bout.

        To test this, she \textbf{runs an online machine learning model} to estimate
        latent variables from prefrontal cortex activity, predicting when this
        peak will occur. She then \textbf{optogenetically inactivates the neural
        population} at the forecasted time.

        Because inactivation prevented the mouse from initiating a foraging bout,
        her hypothesis was supported.
    }

    \only<5>{
    Notes:

    \begin{enumerate}

        \item for a week-long experiment, the size of the behavioural and neural
            recordings exceeds 200~terabytes. Due to the large datasets sizes,
            we will \textbf{bring users to data}, instead of data to users.
            Data will be stored in the cloud, computation will run on the cloud
            and user computers will only display small data and analysis
            results.

        \item we will \textbf{share data, hardware specifications and
            open-source software openly}.

    \end{enumerate}
    }

\end{frame}

\end{document}
