% # Context

% ## RTML is used in many disciplines

Real time machine learning (RTML) in commonly used pervasively across
disciplines.
%
For example, in climate and environmental monitoring it is used for real-time
flood or wildfire detection from satellite or sensor data, and for predicting
extreme weather events using streaming radar and temperature data.
%
Another example is food delivery where RTML is used for predicting delivery
time based on live traffic, restaurant queues, and historical patterns and
dynamic route planning for drivers and shoppers.

For instance, Uber Eat selects the
best person to deliver your food, the best place to buy the food, the best
route to deliver your food, based on advanced analysis of real-time traffic
information, drivers information and food availabily.

Real time weather forecassting.

Real time trading.

Real time best route information.

% ## Despite its potential in neuroscience, it has not find many applications

Conventional neuroscience data analysis is most often performed offline. An
experiment is performed, experimental data is saved to disk, data is analyzed
and results are published.

A new form of online data analysis if possible, where data is analyzed while it
is being collected, real-time analysis results determine how the experiment
unfolds (e.g., what visual/auditory/odour stimuli to present to experimental
subjects, how to guide neural modulations) and what part of the recorded data
to save and what to discard.

% Bonsai & Bonsai.ML

% Need to build adaptive implementations of real-time machine learning algorithms

% Approach

Requirements for real time ML
- hard time constraints
- on pass algorithms

% Deliverables
%   - repository of real time ML methods for neuroscience experimental control
%     adapted to work in non-stationary environments
%
%   - collaborate with scientists at the SWC and AIND to apply the RTML methods
%     to neuroscience problems
