\subsection{Deliverables}

\begin{enumerate}

    \item hardware specifications for long-duration recordings of behaviour and
        neural activity (foraging arenas, video cameras, ultrasound
        microphones, weight scales, Neuropixels probes, commutator,
        \href{https://www.nature.com/articles/s41592-024-02521-1}{ONIX})

    \item software for managing long-duration recordings (data storage, data
        indexing)

    \item software for online/offline quality control

    \item software for online data visualisation

    \item software for online data analysis

    \item software for intelligent close-loop control of experiments

\end{enumerate}

\subsection{Previous work}

\begin{itemize}

    \item  at the SWC we have performed foraging experiments

    \begin{itemize}

        \item lasting xx weeks and recording behaviour only
        \item lasting yy weeks and recording behaviour and electrophysiology
        \item data is stored in files and in a MySQL database

    \end{itemize}

    \item  items 1--3: above have been completed for the SWC foraging experiments

    \item  item 4: we have developed some online data visualisation tools in Bonsai

    \item  items 5 and 6: funded by BBSRC, we have integrated into Bonsai tools for online data
  analysis, to be used for the close-loop control of experiments

    \begin{itemize}

        \item estimate kinematics of mice

        \item estimate kinematic states of mice using Hidden Markov Models

        \item clusterless point-process decoder of mice position and replay from spikes

    \end{itemize}

    \item  disseminated detailed documentation on software used at the SWC to control
        NaLoDuCo experiments (see
        \href{https://github.com/SainsburyWellcomeCentre/aeon_mecha.}{repo})

    \item  disseminated detailed documentation on machine learning methods integrated
  into Bonsai for analyzing behavioral and neural tie series in real time (see
        \href{https://bonsai-rx.org/machinelearning/.}{repo})

\end{itemize}

\subsection{Future work}

\begin{itemize}

    \item item 3: at the SWC and at the AIND we have developed several tools for offline quality control. We next need to build online versions of them.

    \item item 4: develop more software for data visualisation.

    \item items 5 and 6: develop more software for online data analysis.

    \begin{itemize}

        \item online estimate of latent variables from Neuropixels recordings.

        \item online estimate of RL models.

    \end{itemize}

    \item help Dr.~Carl Schoonover (AIND) use hardware and software developed at the SWC/GCNU to create his olfaction NaLoDuCo experiments.

\end{itemize}
