
\subsection{Non-Stationarity}

Most current data analysis methods assumme some sort of stationarity in the
data used to estimate and evaluate their models. This assumption may be
reasonable for experimental data from short duration experiments. However, for
long-duration experiments, with subjects learning and adapting over time, this
stationarity assumption is questionable~\citep{ditzlerEtAl15}.

Here we will rigorously test data stationarity and adapt the proposed methods
to work in non-stationary evnvironments, when necessary.

\subsection{Deliverables}

\begin{enumerate}

    \item methods to analyse behavioural and electrophysiological recordings
    from NaLoDuCo experiments that are \textbf{online} and \textbf{adaptive to
    non-stationarity} in measurements.

    For behavioural data, we will investigate methods to

    \begin{itemize}

        \item track multiple body parts of animals (deep neural networks)

        \item infer kinematics of foraging mice (linear dynamical systems)

        \item segment behaviour into discrete states

        \item characterize short- and long-term periodicities in behavior

        \item infer the rules that govern mice behaviour from behavioural
        observations only] (i.e., policy inference).

    \end{itemize}

    For neural data, we will investigate methods to:
    \begin{itemize}

        \item estimate low-dimensional continual representations of neural
        activity (i.e., latents inference)

        \item segment neural activity into discrete states

        \item characterize short- and long-term periodicities in neural
            activity

        \item decode environment variables from neural activity

    \end{itemize}

    \item integration of these methods into DANDI, as in Dendro, so that users
    can run them on NaLoDuCo datasets stored in DANDI.

\end{enumerate}

\subsection{Previous work}

\begin{itemize}

    \item At the Gatsby we have invented several methods for the
        characterisation of neural time
        series~\citep[e.g.,][]{yuEtAl09,dunckerAndSahani18,ruttenEtAl20,yuEtAl24,buesingEtAl12a,buesingEtAl12b,mackeEtAl15,soulatEtAl21,walkerEtAl23,turnerAndSahani14,osheaEtAl22}

    \item \href{https://github.com/magland/dendro}{Dendro} allows to perform
        advanced data analysis on DANDI. It allows to reuse previous analysis.

\end{itemize}
